Radiocarbon measurement of CO~2~ from carbonate minerals using a gas ion source with an open split interface

### 2022-03-18 doc/AMS-15/AMS-15_paper/AMS-15_paper.Rmd
Expand figure captions, reorganize results.

Determination of ideal flows.
Testing is needed to determine minimum sample flow to avoid atmosphere entering the split. TODO: move flow ratio test to results.
## Method development

@@

A suite of carbonate macrofossils consisting of shell and coral fragements from a series of coastal marine cores REF were measured as graphite and as CO2 using the HGIS method.

@@

Quantification of method

Cost vs quality discussion

ideas for further development of method

@@

Fig. 1. Performance of the NOSAMS MCSNICS hybrid gas ion source.

Fig. 2. Assessment of blanks. (a) Current dependent blank. (b) Fraction modern of large, radiocarbon dead samples compared to sample current. Blank rises with decreasing current, as expected, but may be treated as a single blank with a simple correction for full-size samples producing greater than XX uA C- current.

Fig. 3. Relationship of sample size to C- ion current. There is no strong relationship between size and current for samples larger than XX ug C.

Fig 4. Agreement of standards with graphite consensus values.

Fig 5. Comparison of Haiti core samples measured as graphite.

>>>>

Determination of ideal flows.
Testing is needed to determine minimum sample flow to avoid atmosphere entering the split. TODO: move flow ratio test to results.

## Method development

@@

A suite of 15 carbonate macrofossils consisting of shell and coral fragements from a series of coastal marine cores REF were measured as graphite and as CO2 using the HGIS method to further test the method and serve to demonstrate a potential application.
coral macrofossils from a subtropical coastal marine sediment core using our gas 
method and compared to the same samples measured using the traditional 
graphite method.

To further test the method and a potential application, we measured 15 shell and 
coral macrofossils from a subtropical coastal marine sediment core using our gas 
method and compared to the same samples measured using the traditional 
graphite method. 
We found the HGIS and graphite measurements to be in good agreement, with no 
significant bias (figure 6). Mean per-sample error was 1 pMC for HGIS measurements, 
compared with 0.3 pMC for graphite.
The accuracy of age-depth models is often driven by environmental factors and 
sparse data, rather than the precision of individual dates [3]. The lower time and 
cost of the hybrid method allows analysis of more samples. This may produce more 
accurate age models, despite reduced precision of individual measurements.

@@

## Quantification of method

## Cost vs quality discussion

The accuracy of age-depth models is often driven by environmental factors and 
sparse data, rather than the precision of individual dates [3]. The lower time and 
cost of the hybrid method allows analysis of more samples. This may produce more 
accurate age models, despite reduced precision of individual measurements.

## ideas for further development of method

@@

Fig. 1. Schematic of the open split inlet. Two valco valves provide the ability to admit helium and live or dead tank standards. The double needle  is housed in a VICI cross and allows displacement helium to flow into the sample vial and the sample-helium mixture to flow to the open split, where a capillary delivers the mixture to the ion source.

Fig. 2. Performance of the NOSAMS MCSNICS hybrid gas ion source. a. Modelled flows of helium and CO2 to the ion source during displacement of a pure CO2 vial sample with helium. b. Measured 12C- ion current during the same dillution experiment. c. Measured ion current for modelled CO2 flow. The stability of the source over a range of flow produces stable current for up to 40 m of measurement time.

Fig. 3. Relationship of sample size to C- ion current. There is no strong relationship between size and current for samples larger than 800 ug C.

Fig. 4. Assessment of blanks. (a) Current dependent blank model and data. Fm and current of measurements from dillution tests of live and dead gas are used to generate linear fits which should cross at the Fm and current of the constant current blank. Direct determination of Fm and current of the blank are also shown. (b) Fraction modern of large, radiocarbon dead samples compared to sample current. Blank rises with decreasing current, as expected, but may be treated as a single blank with a simple correction for full-size samples producing greater than 3 uA C- current.

Fig 5. Agreement of standards with graphite consensus values. Difference of HGIS measurements from consensus for individual samples are plotted against their expected value. Expected values are spread to avoid overlapping points. Error bars represent one sigma sample errors. The black line is the mean difference from expected value for all measurements, and the blue lines are one standard deviation of all measurements.

Fig 6. Comparison of Haiti core samples measured as graphite. a. Difference between HGIS and graphite measurements of the same materials vs. core depth (should this be mean of values?). Samples are offset on the horizontal axis for clarity. Error bars are measurement error for each sample. b) comparison of age-depth models produced using HGIS and graphite measurements.
